---
title: "Capstone"
author: "Ben Bartoszek"
date: "11/4/2019"
output:
  word_document: default
  html_document: default
---

#Load the rstudio packages
```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(readxl)
library(ggplot2)
library(reshape2)
library(ggthemes)
library("RColorBrewer")
library(scales)
library(wesanderson)
library(car)
library(texreg)
```

#Load the data and name each data frame
```{r}
Health_Deficiencies <- read.csv("HealthDeficiencies_Download.csv")
Fire_Safety_Deficiencies <- read.csv("FireSafetyDeficiencies_Download.csv")
Health_Providers <- read.csv("ProviderInfo_Download.csv")
Health_Providers_18 <- read.csv("ProviderInfo_Download_2018.csv")
Health_Providers_17 <- read.csv("ProviderInfo_Download_2017.csv")
CMS_Regions <- read.csv("CMS_Regions.csv")
Penalties <- read.csv("Penalties_Download.csv")
```

#Summarize the data frame
```{r}
summary(Health_Deficiencies)
```


#Combine the Fire Stafety Deficiencies data frame with the Health Deficiencies data frame
```{r}
Health_Deficiencies <- rbind(Health_Deficiencies, Fire_Safety_Deficiencies)
```

#Convert pnlty_date to a date format, extracts the year and creates a new column with the year in the Penalties data frame
```{r}
Penalties$pnlty_date <- as.Date(Penalties$pnlty_date, "%Y-%m-%d")
Penalties$Year <- lubridate::year(Penalties$pnlty_date)
```

#Rename provnum to PROVNUM to match the other data frames
```{r}
Penalties <- Penalties %>% rename(PROVNUM = provnum)
```

#Create a new data frame, rename three of the columns to match other data frames and reshape PROVNUM and Year as characters
```{r}
Penalties_To_Join <- data.frame(Penalties$PROVNUM, Penalties$Year, Penalties$fine_amt)
Penalties_To_Join <- rename(Penalties_To_Join, PROVNUM = Penalties.PROVNUM, Year = Penalties.Year, fine_amt = Penalties.fine_amt)
Penalties_To_Join$PROVNUM <- as.character(Penalties_To_Join$PROVNUM)
Penalties_To_Join$Year <- as.character(Penalties_To_Join$Year)
```

#Combine address columns to create a full street address
```{r}
Health_Deficiencies <- unite(Health_Deficiencies, "Full_Address", ADDRESS, CITY, STATE, ZIP, sep = ", ", remove = FALSE)
Health_Providers <- unite(Health_Providers, "Full_Address", ADDRESS, CITY, STATE, ZIP, sep = ", ", remove = FALSE)
Health_Providers_17 <- unite(Health_Providers_17, "Full_Address", ADDRESS, CITY, STATE, ZIP, sep = ", ", remove = FALSE)
Health_Providers_18 <- unite(Health_Providers_18, "Full_Address", ADDRESS, CITY, STATE, ZIP, sep = ", ", remove = FALSE)
```

#Create data frames that includes the provider information for the past three years and rename the columns to match the other data frames
```{r}
Health_Providers_2017 <- data.frame(Health_Providers_17$provnum, Health_Providers_17$PROVNAME, Health_Providers_17$Full_Address, Year = "2017", Health_Providers_17$OWNERSHIP, Health_Providers_17$BEDCERT, Health_Providers_17$overall_rating, Health_Providers_17$quality_rating, Health_Providers_17$staffing_rating, Health_Providers_17$survey_rating, Health_Providers_17$STATE)
Health_Providers_2017 <- rename(Health_Providers_2017, PROVNUM = Health_Providers_17.provnum, PROVNAME = Health_Providers_17.PROVNAME, Full_Address = Health_Providers_17.Full_Address, OWNERSHIP = Health_Providers_17.OWNERSHIP, BEDCERT = Health_Providers_17.BEDCERT, Overall_Rating = Health_Providers_17.overall_rating, Quality_Rating = Health_Providers_17.quality_rating, Staffing_Rating = Health_Providers_17.staffing_rating, Survey_Rating = Health_Providers_17.survey_rating, STATE = Health_Providers_17.STATE)
Health_Providers_2018 <- data.frame(Health_Providers_18$PROVNUM, Health_Providers_18$PROVNAME, Health_Providers_18$Full_Address, Year = "2018", Health_Providers_18$OWNERSHIP, Health_Providers_18$BEDCERT, Health_Providers_18$OVERALL_RATING, Health_Providers_18$QUALITY_RATING, Health_Providers_18$STAFFING_RATING, Health_Providers_18$SURVEY_RATING, Health_Providers_18$STATE)
Health_Providers_2018 <- rename(Health_Providers_2018, PROVNUM = Health_Providers_18.PROVNUM, PROVNAME = Health_Providers_18.PROVNAME, Full_Address = Health_Providers_18.Full_Address, OWNERSHIP = Health_Providers_18.OWNERSHIP, BEDCERT = Health_Providers_18.BEDCERT, Overall_Rating = Health_Providers_18.OVERALL_RATING, Quality_Rating = Health_Providers_18.QUALITY_RATING, Staffing_Rating = Health_Providers_18.STAFFING_RATING, Survey_Rating = Health_Providers_18.SURVEY_RATING, STATE = Health_Providers_18.STATE)
Health_Providers_2019 <- data.frame(Health_Providers$PROVNUM, Health_Providers$PROVNAME, Health_Providers$Full_Address, Year = "2019", Health_Providers$OWNERSHIP, Health_Providers$BEDCERT, Health_Providers$Overall_Rating, Health_Providers$Quality_Rating, Health_Providers$Staffing_Rating, Health_Providers$SURVEY_RATING, Health_Providers$STATE)
Health_Providers_2019 <- rename(Health_Providers_2019, PROVNUM = Health_Providers.PROVNUM, PROVNAME = Health_Providers.PROVNAME, Full_Address = Health_Providers.Full_Address, OWNERSHIP = Health_Providers.OWNERSHIP, BEDCERT = Health_Providers.BEDCERT, Overall_Rating = Health_Providers.Overall_Rating, Quality_Rating = Health_Providers.Quality_Rating, Staffing_Rating = Health_Providers.Staffing_Rating, Survey_Rating = Health_Providers.SURVEY_RATING, STATE = Health_Providers.STATE)
```

#Combine the three data frames containing provider information for the past three years
```{r}
Health_Providers_All <- rbind(Health_Providers_2019, Health_Providers_2018, Health_Providers_2017)
```

#Create a new column that identifies each deficiency as "Substandard Care" or "Not Substandard Care" based off of the scope of the deficiency
```{r}
Health_Deficiencies$Substandard_Care <- ifelse(Health_Deficiencies$SCOPE == "F"| Health_Deficiencies$SCOPE == "H" | Health_Deficiencies$SCOPE == "I" | Health_Deficiencies$SCOPE == "J" | Health_Deficiencies$SCOPE == "K" | Health_Deficiencies$SCOPE == "L", "Substandard Care", "Not Substandard Care")
```

#Join the CMS_Regions data frame with the Health_Providers_All data frame by state
```{r}
Health_Providers_All <- left_join(Health_Providers_All, CMS_Regions, by = "STATE")
```

#Create new column with a simplified ownership type for each provider to narrow down the ownership categories for the regression model
```{r}
Health_Providers_All$Own_Type <- ifelse(Health_Providers_All$OWNERSHIP == "For profit - Corporation"| Health_Providers_All$OWNERSHIP == "For profit - Individual" | Health_Providers_All$OWNERSHIP == "For profit - Limited Liability company" | Health_Providers_All$OWNERSHIP == "For profit - Partnership", "For Profit", ifelse(Health_Providers_All$OWNERSHIP == "Government - City"| Health_Providers_All$OWNERSHIP == "Government - City/county" | Health_Providers_All$OWNERSHIP == "Government - County" | Health_Providers_All$OWNERSHIP == "Government - Federal" | Health_Providers_All$OWNERSHIP == "Government - Hospital district" | Health_Providers_All$OWNERSHIP == "Government - State", "Government", ifelse(Health_Providers_All$OWNERSHIP == "Non profit - Church related"| Health_Providers_All$OWNERSHIP == "Non profit - Corporation" | Health_Providers_All$OWNERSHIP == "Non profit - Other", "Non Profit", "N/A")))
```

#Create new column in the Health_Deficiencies data frame with the year that the Deficiency was received based on the cycle
```{r}
Health_Deficiencies$Year <-  ifelse(Health_Deficiencies$cycle == "1", "2019", ifelse(Health_Deficiencies$cycle == "2", "2018", ifelse(Health_Deficiencies$cycle == "3", "2017", "N/A" )))
```

#Create a column in the Health_Deficiencies data frame with a 1 to count the number of deficiencies by provider
```{r}
Health_Deficiencies$Count <- "1"
```

#Create a new data frame with the count of the number of deficiencies for each PROVNUM and Year for each category of "Substandard Care" and "Not Substandard Care"
```{r}
HD_to_melt <- Health_Deficiencies %>% select(PROVNUM, Year, Substandard_Care, Count)
HD_Melt <- melt(HD_to_melt, id.vars = c("PROVNUM", "Year", "Substandard_Care"), variable.name = "Count")
HD_dcase <- dcast(HD_Melt, PROVNUM + Year ~ Substandard_Care, value.var = c("Count"))
```

#Join the HD_dcase data frame with the Health_Providers_All data frame by PROVNUM and Year
```{r}
All_Provider_Deficiencies <- left_join(Health_Providers_All, HD_dcase, by = c("PROVNUM", "Year"))
```

#Replace all N/As in Not Substandard Care and Substandard Care columns with 0
```{r}
All_Provider_Deficiencies$`Not Substandard Care`[is.na(All_Provider_Deficiencies$`Not Substandard Care`)] <- 0
All_Provider_Deficiencies$`Substandard Care`[is.na(All_Provider_Deficiencies$`Substandard Care`)] <- 0
```

#Join the Penalties_To_Join data frame with the All_Provider_Deficiencies data frame by PROVNUM and Year and replace all N/As in fine_amt column with 0
```{r}
All_Provider_Deficiencies <- left_join(All_Provider_Deficiencies, Penalties_To_Join, by = c("PROVNUM", "Year"))
All_Provider_Deficiencies$fine_amt[is.na(All_Provider_Deficiencies$fine_amt)] <- 0
```

#Create linear regression model with fine_amt as the intercept and sumarize the model
```{r}
Model1 <- lm(formula = All_Provider_Deficiencies$fine_amt ~ All_Provider_Deficiencies$Quality_Rating + All_Provider_Deficiencies$Staffing_Rating + All_Provider_Deficiencies$Overall_Rating + All_Provider_Deficiencies$'Not Substandard Care' + All_Provider_Deficiencies$'Substandard Care' + All_Provider_Deficiencies$BEDCERT, data = All_Provider_Deficiencies)
summary(Model1)
```

#Create data frame with only rows that have either one or more Not Substandard Care deficiency or one or more Substandard Care deficiency
```{r}
DF_With_Defs <- All_Provider_Deficiencies[All_Provider_Deficiencies$`Not Substandard Care` > 0 | All_Provider_Deficiencies$`Substandard Care` > 0,]
view(DF_With_Defs)
```

#Create data frame only with rows with a fine amount greater than zero
```{r}
DF_With_Fines <- DF_With_Defs[DF_With_Defs$fine_amt > 0,]
View(DF_With_Fines)
```

#Create linear regression model using a logarithmic transformation on the intercept and sumarize the model
```{r}
Model2 <- lm(formula = log(DF_With_Fines$fine_amt) ~ Overall_Rating + `Not Substandard Care` + `Substandard Care` + BEDCERT + Year + OWNERSHIP + Quality_Rating + Staffing_Rating, data = DF_With_Fines)
summary(Model2)
```

#Create linear regression model using a logarithmic transformation on the intercept and on the BEDCERT coefficient and sumarize the model
```{r}
Model3 <- lm(formula = log(DF_With_Fines$fine_amt) ~ `Not Substandard Care` + `Substandard Care` + log(BEDCERT) + Year + Own_Type + Overall_Rating + Region, data = DF_With_Fines)
summary(Model3)
```

#Create diagnostic plots to analyze the linear regression model
```{r}
plot(Model3)
```

#Calculate the variance inflation factors for the linear regression model
```{r}
vif(Model3)
```

#Create a bar chart with fine_amt against Year and faceted on Region
```{r}
ggplot(data=subset(DF_With_Fines, !is.na(STATE)), aes(x = Year, y = fine_amt)) +
  geom_col(aes(col = Year, fill = Year)) +
  facet_wrap(. ~ Region, scales = "free_x") +
  scale_y_continuous(labels = comma)
```

#Create a bar chart with average fine_amt against Year and faceted on Region
```{r}
ggplot(data=subset(DF_With_Fines, !is.na(STATE))) +
  stat_summary(mapping = aes(x = Year, y = fine_amt, fill = Year), fun.y = "mean", geom = "bar", na.rm = TRUE, inherit.aes = FALSE) +
  facet_wrap(. ~ Region, scales = "free_x") +
  scale_y_continuous(labels = comma)
```

#Create a jittered plot graph with fine_amount against Substandard Care and faceted on Year with a smoothed conditional mean line
```{r}
ggplot(DF_With_Fines, aes(`Substandard Care`, fine_amt)) +
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  facet_wrap(. ~ Year) +
  scale_y_continuous(labels = scales::comma)
```

#Create a jittered plot graph with fine_amount against BEDCERT and faceted on Year with a smoothed conditional mean line
```{r}
ggplot(DF_With_Fines, aes(BEDCERT, fine_amt)) +
  geom_point() +
  geom_jitter() +
  geom_smooth() +
  facet_wrap(. ~ Year) +
  scale_y_continuous(labels = scales::comma)
```

